Stack Safety is a Security Property

### 2021-04-10 tex/main.tex
Polishing and comments

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, with consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}
\fi
%
The implementation and configuration of these  stack-safety mechanisms can be complex,
and often involves compromises trading protection for efficiency.
Thus it would be useful to have a precise formal specification for stack
safety, both as a basis for comparing the security guarantees claimed by different enforcement
techniques and for verifying that these claims are met by particular implementations.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
In essence, memory locations within each stack frame are tagged with the identity of
the function that owns the frame, and access to those locations is then gated
according to the function that is currently executing. Many alternatives can
be implemented on top of this basic idea, by choosing how to identify functions
and deciding just which accesses are checked.  These alternatives vary in their
efficiency, but also in the spatial and temporal precision of their security guarantees.
For example, the \emph{Depth Isolation} policy strictly requires that a function owns any
frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
On the other hand, the \emph{Lazy Tagging and Clearing} policy improves performance by
allowing illicit writes to the frame so long as the written locations are not
subsequently read. The intuition justifying laziness is that any genuinely dangerous writes will ultimately be
detected before they can affect program behavior.  But we clearly need a more formal foundation for
expressing, comparing and verifying these micro-policy security claims, as well as those of
other enforcement technologies.

@@

Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce this property either, as we show in
testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce observational integrity,
at some performance cost.\apt{but we're not testing that, right?}

@@

Our initial description of the properties assumes single simple stack with no sharing between callers and
callees.  We expand it to allow passing of scalar stack data, granting a callee
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)

>>>>

The call stack is a perennial target for low-level attacks
based on memory-safety vulnerabilities, leading to consequences
from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a profusion of
software and hardware protections have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b,Tsampas+19,Georges+21},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\bcp{Yes, going back to MIT days---we should
  include several more of these, if only to give readers the impression that
this is a well-studied mechanism (so formalizing its protections is
useful).} 
\fi
%
The protections offered by such mechanisms are generally described in terms
of a set of concrete examples of attacks that they can prevent, leaving a
more general characterization to the reader's intuition.
Moreover, the details of these mechanisms can be quite intricate
and often involve trading off protection and efficiency.
Thus, it would be useful to have a precise, generic, and formal
specification for stack
safety, both as a basis for comparing the security claims of different
enforcement techniques and for validating more rigorously that these claims
are met by particular implementations.

For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Intuitively, memory cells within each stack frame are tagged with the
identity of
the function that owns the frame, and access to those locations is then gated
according to the function that is currently executing. Roessler and Dehon
explored a number of versions
of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed.  These alternatives
vary in their
efficiency and in the spatial and temporal precision of their security guarantees.
For example, their \emph{Depth Isolation} policy demands that a function
owns any
frame into which it writes; this seems simple and intuitively correct, but can be expensive to implement.
Conversely, their \emph{Lazy Tagging and Clearing} policy improves
performance by
allowing ``illicit'' writes to out-of-scope stack frames so long as the
written locations are not
subsequently read. The justification for this form of laziness is that any
genuinely dangerous writes will ultimately be
detected before they can affect program behavior. \bcp{Awkward transition here.} But we clearly need a
more formal foundation for
expressing, comparing and verifying these micro-policy security claims, as well as those of
other enforcement technologies.

@@

Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce this property either, as we show in
testing~\apt{really??}, but formalizing the ideal property allows us to see how it falls short.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce observational integrity,
at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
pretty much done}

@@

Our initial description of the properties assumes single simple stack with no sharing between callers and
callees.  We expand it to allow passing of scalar stack data, granting a callee
access to the caller's data. We do not model more sophisticated sharing, because to do
so realistically would require a separate protection scheme, such as a capability
model. While this is quite feasible in a tag-based enforcement mechanism, it is properly
a separate policy outside the scope of stack safety. (A tag-based approach can easily
compose multiple such policies.)
\rb{So no pointer passing? Pass by reference is now discussed at the end of
Section 3, however it looks like things get tricky rather quickly.}
